Staged Abstract Interpreters

### 2019-09-07 paper/60_cases.tex
revision

We then integrate the timestamp into our monad stack by adding a staged
@StateT@ monad. The result type is accordingly updated to contain a set of
tuples, where the time is added.
\begin{lstlisting}
  def tick(e: Expr): AnsM[Unit] = for {
    τ <- get_time; u <- liftM(StateTMonad.put((e::τ).take($k$)))
  } yield u
  def alloc(x: String): AnsM[Addr] = for {
    τ <- get_time
  } yield unit(KCFAAddr(x, τ)) // we use unit to turn a current-stage constant to next-stage value
\end{lstlisting}

Every time we call the @eval@ function, we refresh the timestamp by
calling the @tick@ function, which updates the timestamp in the state monad.
The @tick@ function is implemented as prepending the current control expression
being evaluated (@e@) to the existing calling context, and then taking the
first $k$ elements from the list.  The type @Config@ used in caching is also
changed to include timestamps.

@@

A commonly-used technique to improve the running time in control-flow analysis
is store-widening. As shown by \cite{Darais:2015:GTM:2814270.2814308,
DBLP:journals/pacmpl/DaraisLNH17}, store-widening analysis can be easily
implemented in monadic abstract interpreter by swapping the @StateT@ (for store)
and the @SetT@ (for nondeterminism). In this section, we show this optimization
is orthogonal to staging and we just need to similarly swap the same \textit{staged}
version of monads. After the adjustment, we have the following monad stack:
\begin{lstlisting}
  type AnsM[T] = RepReaderT[RepSetT[RepStateT[
                   RepReaderT[RepStateT[IdM, Cache, ?], Cache, ?], Store, ?], ?], Env, T]
\end{lstlisting}

@@

Abstract garbage collection is a technique to reclaim unreachable addresses in
the store while doing the abstract interpretation
\cite{Might:2006:IFA:1159803.1159807}. \citet{DBLP:journals/pacmpl/DaraisLNH17}
showed that, for big-step monadic abstract interpreters, we may add a @Reader@
monad to track a set of root addresses and compute the reachable addresses every
time when we obtain a value. Here, we adopt the idea with staged monads:
\begin{lstlisting}
  type AnsM[T] = RepReaderT[RepReaderT[
                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Store, ?],
                 Set[Addr], ?], Env, T] //add a reader monad for Set[Addr]
\end{lstlisting}

The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We
update the @fix@ function with caching (from Section \ref{sai}) to @fix_gc@,
but the caching part is omitted in the presented code. Several auxiliary
functions are used: @ask_root@ retrieves the current root addresses from the
monad stack; @root@ computes the reachable addresses from a value; and @gc@
performs garbage collection according to the reachable addresses and returns a new
store. For the generic interpreter, we also need to compute the root addresses
for compound expressions, readers may refer to
\cite{DBLP:journals/pacmpl/DaraisLNH17}.

>>>>

We then integrate the timestamp into our monad stack by adding a staged
@StateT@ monad. The result type is accordingly updated to contain the set of
tuples, where the time is added.
\begin{lstlisting}
  def tick(e: Expr): AnsM[Unit] = for {
    τ <- get_time; u <- liftM(StateTMonad.put((e::τ).take($k$)))
  } yield u
  def alloc(x: String): AnsM[Addr] = for {
    τ <- get_time
  } yield unit(KCFAAddr(x, τ)) // we use unit to turn a current-stage constant to next-stage value
\end{lstlisting}

Every time when we call the @eval@ function, we also refresh the timestamp by
calling the @tick@ function, which updates the timestamp in the state monad.
The @tick@ function is implemented as prepending the current control expression
being evaluated (@e@) to the existing calling context, and then taking the
first $k$ elements from the list. The type @Config@ used in the caching
algorithm is also changed to include timestamps.

@@

A commonly-used technique to improve the running time of control-flow analyses
is store-widening. As shown by \cite{Darais:2015:GTM:2814270.2814308,
DBLP:journals/pacmpl/DaraisLNH17}, a store-widening analysis can be easily
implemented in the monadic abstract interpreter by swapping the @StateT@ and
the @SetT@ monad transformer. In this section, we show this optimization is
orthogonal to staging and we just need to similarly swap the same
\textit{staged} version of monads. After the adjustment, conceptually, we have
the following monad stack:
\begin{lstlisting}
  type AnsM[T] = RepReaderT[RepSetT[RepStateT[
                   RepReaderT[RepStateT[IdM, Cache, ?], Cache, ?], Store, ?], ?], Env, T]
\end{lstlisting}

@@

Abstract garbage collection is a technique to reclaim unreachable addresses in
the store while doing the abstract interpretation
\cite{Might:2006:IFA:1159803.1159807}. \citet{DBLP:journals/pacmpl/DaraisLNH17}
showed that, for big-step monadic abstract interpreters, we may add a @Reader@
monad to track the set of root addresses and compute the reachable addresses every
time when we obtain a value. Here, we adopt the idea with staged monads:
\begin{lstlisting}
  type AnsM[T] = RepReaderT[RepReaderT[
                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Store, ?],
                 Set[Addr], ?], Env, T] //add a reader monad for Set[Addr]
\end{lstlisting}

The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We
update the @fix@ function with caching (from Section \ref{sai}) to @fix_gc@
(the caching part is omitted in the presented code snippet). Several auxiliary
functions are used: @ask_root@ retrieves the current root addresses from the
monad stack; @root@ computes the reachable addresses given a value; and @gc@
performs garbage collection according to the given reachable addresses and
returns a new store. For the generic interpreter, we also need to compute the
root addresses for compound expressions; for the detail of this part, readers
may refer to \cite{DBLP:journals/pacmpl/DaraisLNH17}.
